---
title: "R Notebook"
output: html_notebook
---

```{r,warning=FALSE,message=FALSE}
library(dplyr)
library(lubridate)
library(rstan)

```

```{r}
df <- 
  read.csv("NY_cases.csv") 
df<-df%>%select(state,submission_date,new_case, tot_cases)
```

```{r}
lastday <- df %>% 
  mutate(days_away=rev(row_number())-1,
         weight=dgamma(days_away,shape=5.9^2/3.9,rate=5.9/3.9))
  
```

```{r, comment="", echo = FALSE}
writeLines(readLines("reproduction_num.stan"))
```


```{r}
overallinfectivity=sum(lastday$new_case*lastday$weight)
It=lastday$new_case[nrow(lastday)]
a=9/8
b=3/4
```


Approach:

step1: Calculate daily R_t (the estimated mean of floating weekly transimission rate) in New York State with function estimate_R from package EpiEstim. We choose to specify the distribution of the serial period and assumes it follows a gamma distribution with mean of 5.9 and standard deviation of 3.9, according to the paper of Liu et al.. 


```{r, Stan, results = "hide", message = FALSE, warning = FALSE}
post <- stan("reproduction_num.stan",   
             data = list(infect=overallinfectivity,I=It, prior_only = 0, alpha = a, beta=b))
```

```{r}
post
```



## Posterior Planes

```{r, fig.width=10, fig.height=5, small.mar = TRUE}
pairs(post, pars = "p", include = FALSE) 
```
```{r, fig.width=10, fig.height=5, small.mar = TRUE}
draws <- as.data.frame(post) %>% select(-starts_with("p"))
mean(draws$Rt)
```
Essentially we can compute the Rt for all time points in the dataset,assuming the same serial interval distribution. 

```{r, loop, results = "hide", message = FALSE, warning = FALSE}
for (i in 2:nrow(df)){
  dt<-df[1:i,] %>% 
  mutate(days_away=rev(row_number())-1,
         weight=dgamma(days_away,shape=5.9^2/3.9,rate=5.9/3.9))
  overall=sum(dt$new_case*dt$weight)
  I_t=dt$new_case[nrow(dt)]
  a=9/8
  b=3/4
  posterior <- stan("reproduction_num.stan",   
             data = list(infect=overall,I=I_t, prior_only = 0, alpha = a, beta=b))
post.df <- as.data.frame(posterior) %>% select(-starts_with("p"))
df$Rt[i]<-mean(post.df$Rt)
}


```



