---
title: "Bayesian Inference Final Project"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Introduction

The reproduction number $R_t$ is a common measure of transmissibility of an epidemic disease. By definition, $R_t$ is the average number of secondary cases caused by an infected individual. $R_t$ falling below 1 indicates the disease is unlikely to sustain. The change of $R_t$ across time can be used as a proxy for epidemic trajectories. Following the method proposed by Cori et al., $R_t$ can be calculated by using Bayesian parametric estimation. In this project, I use the same method to compute the $R_t$ of COVID-19 in New York before the presence of effective vaccination (Mar 2020 to Dec 2020). 


```{r,warning=FALSE,message=FALSE}
library(dplyr)
library(lubridate)
library(rstan)

```

```{r}
df <- 
  read.csv("NY_cases.csv") 
df<-df%>%select(state,submission_date,new_case, tot_cases)
```

## Method

Assuming the number of reported incident cases follows a poisson process, we have:

$P(I_t|I0,I1,..,I_{t-1},w_s,R_t)=\frac{(R_t\Lambda_t(w_s))^I_t exp(-R_t\Lambda_t(w_s)}{I_t!}$,

where $I_t$ is the number of incident cases arising at time t and
$\Lambda_t(w_s)$ is the overall infectivity and can be computed as:

$\Lambda_t(w_s)=\sum_{s=1}^{t}{I_{t-s}w_s}$.

$w_s$ is the serial interval distribution, which is the time between the onset of symptoms in a primary case and he onset of symptoms in secondary cases. It serves as the weight for reported new cases at each time step before time t. 
I choose the distribution of the serial period of COVID-19 follows a gamma distribution with mean of 5.9 and standard deviation of 3.9, according to the paper of Liu et al.. I also choose the prior of $R_t$ to follow a gamma distribution with mean = 1.5 and standard deviation of 2.

With the above assumptions and procedure, I use Stan to yield the posterior distribution of $R_t$ on the last day of 2020 (Dec 31st, 2020) with all previous reported cases.
 
```{r}
#get the weight of previous days using a gamma distribution
lastday <- df %>% 
  mutate(days_away=rev(row_number())-1,
         weight=dgamma(days_away,shape=(5.9/3.9)^2,rate=5.9/3.9^2))
  
```


```{r, comment="", echo = FALSE}
#Write stan code for estimation
writeLines(readLines("reproduction_num.stan"))
```


```{r}
#calculating overall infectivity  
overallinfectivity=sum(lastday$new_case*lastday$weight)
#indicate new case on Dec 31,2020
It=lastday$new_case[nrow(lastday)]
#choose prior parameters (shape and rate)
a=9/16
b=3/8
```


```{r, Stan, results = "hide", message = FALSE, warning = FALSE}
post <- stan("reproduction_num.stan",   
             data = list(infect=overallinfectivity,I=It, prior_only = 0, alpha = a, beta=b))
```

```{r}
post
```
The mean of the posterior distribution of $R_t$ on Dec.31st, 2020 in New York is 1.57 with a 0.95 credible interval of (1.54,1.60).


## Posterior Planes

```{r, fig.width=10, fig.height=5, small.mar = TRUE}
pairs(post, pars = "p", include = FALSE) 
```
```{r}
draws <- as.data.frame(post) %>% select(-starts_with("p"))
mean(draws$Rt)
```

## Different Assumptions 

I then modify the approach by using a negative binomial likelihood instead of a poisson likelihood since negative binomial distribution allows for overdispersion. The negative binomial distribution has an extra dispersion parameter $\phi$ and it is assumed to follow a half cauchy distribution with location parameter =5 and scale = 5.

```{r, comment="", echo = FALSE}
#Write stan code for estimation
writeLines(readLines("reproduction_num2.stan"))
```


```{r, results = "hide", message = FALSE, warning = FALSE}
post_neg <- stan("reproduction_num2.stan",   
             data = list(infect=overallinfectivity,I=It, prior_only = 0, alpha = a, beta=b, mu=5,sigma=5))
```
```{r}
post_neg
```

After modification, the mean of the posterior distribution of $R_t$ on Dec.31st, 2020 in New York is 1.70 with a 0.95 credible interval of (0.77,3.41). The negative binomial likelihood gives a wider credible interval which implies higher uncertainty in estimation. 

```{r, fig.width=10, fig.height=5, small.mar = TRUE}
pairs(post_neg, pars = "p", include = FALSE) 
```




Essentially, we can compute the posterior distribution of $R_t$ for all time points in the data set,assuming the same serial interval distribution. Using the mean of the posterior distributions, a graph of $R_t$ can be plotted.

```{r, loop,eval=FALSE}
for (i in 2:nrow(df)){
  dt<-df[1:i,] %>% 
  mutate(days_away=rev(row_number())-1,
         weight=dgamma(days_away,shape=5.9^2/3.9,rate=5.9/3.9))
  overall=sum(dt$new_case*dt$weight)
  I_t=dt$new_case[nrow(dt)]
  a=9/16
  b=3/8
  posterior <- stan("reproduction_num.stan",   
             data = list(infect=overall,I=I_t, prior_only = 0, alpha = a, beta=b))
post.df <- as.data.frame(posterior) %>% select(-starts_with("p"))
df$Rt[i]<-mean(post.df$Rt)
}

```


### Reference

Anne Cori, Neil M. Ferguson, Christophe Fraser, Simon Cauchemez, A New Framework and Software to Estimate Time-Varying Reproduction Numbers During Epidemics, American Journal of Epidemiology, Volume 178, Issue 9, 1 November 2013, Pages 1505–1512, https://doi.org/10.1093/aje/kwt133

Liu, X., Xu, X., Li, G., Xu, X., Sun, Y., Wang, F., Shi, X., Li, X., Xie, G., & Zhang, L. (2021). Differential impact of non-pharmaceutical public health interventions on COVID-19 epidemics in the United States. BMC Public Health, 21(1), 965. https://doi.org/10.1186/s12889-021-10950-2

